### 2024-04-09 paper.tex

\documentclass[onecolumn, 12pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{setspace}

\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{footnote}
\usepackage{endnotes}
\usepackage[bottom, multiple]{footmisc}
\usepackage[level]{fmtcount}

\usepackage{float}
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{ {./graphics} }

\usepackage[style=ieee, backend=biber]{biblatex}
\addbibresource{sources.bib}

\usepackage[pdfusetitle, hidelinks]{hyperref}

\makesavenoteenv{figure}
\SetCiteCommand{\cite}

\title{Sailing the High Seas:\\ A Christian Exploration of Piracy}
\author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
\date{Wheaton College\\\today}

\begin{document}
\maketitle

\section{Introduction}

\subsection{Background Information}

\begin{displayquote}
  \textins{T}hat ideas should freely spread from one to another over the globe,
  for the moral and mutual instruction of man, and improvement of his
  condition, seems to have been peculiarly and benevolently designed by nature,
  when she made them, like fire, expansible over all space, without lessening
  their density at any point, and like the air in which we breathe, move, and
  have our physical being, incapable of confinement or exclusive appropriation.
  Inventions then cannot, in nature, be a subject of property.
  Thomas Jefferson~\cite{barlow:wine}
\end{displayquote}
It is generally accepted that piracy is an illegal activity, but what is often overlooked
is the legitimate role that piracy plays in a variety of contexts. When it comes to
subverting oppressive regimes, for instance, the tools that piracy provides are
indispensable and for protecting the interests of consumers from predatory business
practices, the abilities of piracy are unmatched.

Needless to say, with every unregulated system comes abuse and piracy is no exception.
On top of that, piracy, on an individual level, is extremely difficult to combat because
of the immaterial nature of the items that would need to be protected. Ideas flow freely
between people especially when the Internet is involved, so stopping this flow becomes an
exponentially more difficult problem the more people have the restricted data. The ease
with which ideas propagate lead people like Thomas Jefferson to claim that they are
``incapable of confinement or exclusive appropriation.'' Consequently, two parties form:
those that support piracy for its free and open spread of all information and those that
see piracy as a threat to the further advancement of technology and ideas.

Before proceeding further into the issue of piracy, it is important to have a grasp of
what piracy means insofar as how this paper interprets it. The dictionary defines piracy
as ``the unauthorized use of another's production, invention, or conception especially in
infringement of a copyright.'' While this covers the technical definition of piracy, this
paper elects to utilize a more wholistic definition. Piracy not only includes the
unauthorized use of someone's work, but it includes the tools and techniques that are
commonly used to accomplish this task such as torrent trackers, VPNs, and torrent clients.

As a Christian, how should one approach such a tool and how can it be utilized for the
betterment of God's kingdom? In the paper that follows, piracy will be analyzed through a
number of lenses, culminating in a Christian response to this issue. In doing so, the hope
is that the reader attains a fresh, wholistic understanding of the potential value of
piracy.

\subsection{Thesis/Argument}

Piracy should not be hastily thrown aside.
Piracy is an excellent tool for disseminating information, services, content,
\&c. that may not have been available any other way.

\section{Arguments For Piracy}
As mentioned above, even though piracy is illegal, there is still merit to the tools of
piracy and the values behind them. The merits of piracy can be divided into three realms:
equal access to information, governance, and privacy. 
\subsection{Equal Access}
At its core, piracy is primarily concerned with accessing information of all kinds,
regardless of whether someone is authorized to access that information or not. This is
what piracy is best at and, contrary to popular belief, it is not always a bad thing.

Take the restricted access to academic journal articles, for instance. Many academic
journals require all readers to pay not an insignificant sum to access research articles
that could contain insight that is valuable to the human race as a whole. Because of the
high cost of entry, certain groups are automatically excluded from reading and building
upon the research that has already been done. John Barlow puts the situation eloquently in
\textit{Selling Wine Without Bottles: The Economy of the Mind on the Global Net} when he
says, \textcquote{barlow:wine}{I am not comfortable with a model which will restrict
inquiry to the wealthy.} In the case of academic journal articles, why shouldn't anyone be
able to better themselves by consuming quality research articles? All humans should be
allowed to continue to learn more about God's creation without being bound by their
financial status. Just as how someone can go to the library and checkout any book that
tickles their fancy, they should be freely able to learn from the ongoing research of our
world. Of course one should recognize that not all information is safe or helpful to be
freely available on the internet, but that is a topic for a later section.

A more concrete instance of where paywalls can be directly harmful to the human race is
with regards to medical research. Till et al. stresses how important access to medical
research is in their article \textit{Who is pirating medical literature? A bibliometric
review of 28 million Sci-Hub downloads}. \blockcquote{till:medical-literature} {Access to
the medical literature is essential for both the practice of evidence-based medicine and
meaningful contribution to medical sciences. Nonetheless, only 12\% of newly published
papers are freely accessible online, and, as of 2014, only 3 million of the 26.3 million
articles indexed on PubMed were available on the site's repository of free materials,
PubMed Central. Access to paywall-protected literature remains primarily through
institutional subscriptions. Such subscriptions are costly and many struggle to afford
access. The result is a disparity in access to the medical literature, particularly for
those in low-income and middle-income countries (LMICs).}
According to Till et al. the high subscription cost to access medical journal articles
means that those in less advantaged countries cannot afford access to the material, which
is detrimental to the furthering of medical sciences. In light of this, it should come as
no surprise that \textcquote{till:medical-literature} {Nearly 1 million articles published
by medical journals are downloaded on Sci-Hub each month.} Piracy enables the less
fortunate countries to access valuable medical data that is essential to furthering
medical research for the human race. In such a specific case as this, it should not be
controversial to conclude that piracy is a good thing.

\begin{enumerate}
  \item Piracy allows equal access to information
  \item \textcquote{barlow:wine}{I am not comfortable with a model which will restrict inquiry to the wealthy.}

  \item Increased prices are an artificial barrier to access certain
    information. Piracy provides a circumvention to said artificial barrier.
    Piracy gives consumers a way to fight systems they see as
    unfair/censorship. Piracy is a tool of the consumer to control/balance the
    economy~\cite{darnton:pirating-and-publishing, bohannon:everyone}

  \item People pirate medical content.~\cite{till:medical-literature}
    Medical information is valuable to everyone. Having medical information is
    crucial to advancements in medicine, but sometimes the cost is too high.
    The paywalls, in this case, are directly harmful to ends that are
    beneficial to everyone.
\end{enumerate}

\subsection{Governance}

\begin{enumerate}
  \item Dissemination of information through means of piracy and torrenting
    gives access to those within areas where Internet shutdowns are commonplace.

  \item Piracy provides a way around an oppressive regime, which are on the
    rise.~\cite{current:jigsaw}

  \item IP laws arbitrarily restrict free speech / invention just because
    someone came up with it first.~\cite[1339]{lemley:faith-based} They are
    based on the assumption that, in doing so, they are encouraging creativity
    (to be first to invent/create), when, in reality, creativity doesn't always
    come first, but may rather be an improvement upon the established. The most
    creative solution is not always the first one. IP laws and restrictions
    block slight innovations upon the existing intellectual property.
\end{enumerate}

\subsection{Privacy}

\begin{enumerate}
  \item Piracy gives you access to cracked versions of software that provide
    more sensible privacy/freedom than what is accessible
    legally.~\cite{stallman:right-to-read}
\end{enumerate}

\section{Arguments Against Piracy}
While piracy enables free access to information and content, it is crucial to
faithfully and charitably explore the potential dangers and economic
implications associated with this practice. The arguments against piracy are
not easy, nor should be, easy to dismiss. This paper will focus mainly on the
ethical and economical concerns that the use of piracy introduces into society.

\subsection{Access to Dangerous/Illegal Content}
Piracy has been associated with unrestricted access to information and
content. As previously discussed, this aspect of piracy provides the means for
those in need to access information that would otherwise be unavailable to them;
similarly, piracy can thus provide avenues to bypass governmental restrictions
on public information. This same benefit, however, becomes a strong argument
against piracy when what is distributed is dangerous or illegal content.

\subsubsection{Criminality of the Seeder and Leecher}
This paper has focused on a more wholistic definition that encompasses
the tools, techniques, motivations, and content, but the motivations and
content that are involved in piracy should not be ignored. Piracy does involve
a form of theft and dissemination of an individual's or organization's
intellectual property. Piracy, thus can and most often does violate laws in
most jurisdictions. Piracy, when it uses torrenting, involves both the
seeder and the leecher (or the uploader and downloader, if not using torrenting).
Depending on the legislation of the country in question, the leecher may or
may not be found guilty of a crime, but this is complicated by the global
context of digital piracy.

In the United States, the downloader/streamer/leecher are generally considered
offenders that could be subject of fines or even criminal charges. The severity
of the penalties of piracy in these cases pose a major argument against the use
of piracy, at least in the United States or nations where the individual
downloading can be found guilty of a crime. ``Fair use'' and statutes of
limitations may alleviate penalties, but repeat offences may also lose one's
favor in these cases.~\cite{felonies.org}

\subsubsection{Malware}
Piracy may openly distribute dangerous illegal content, but arguably the
larger threat is the dangerous content that remains undisclosed. Piracy
circumvents mainline distribution channels that may incorporate secure policies
of data integrity, authenticity, and privacy. Some piracy tools provide
some ways to vet the authenticity of an uploader based on past uploads or based
on the user's relationship with the piracy/torrenting
site\footnote{%
  \href{https://pirates-forum.org/Thread-ThePirateBay-Want-Trusted-VIP-Status-Pink-Green-Skull-See-Here}{ThePirateBay},
  for example provides icons to display whether a user is \emph{Trusted},
  \emph{VIP}, or \emph{Moderator}%
}, but not all sites provide these metrics or warnings for the average end-user.
Instead, the burden is often on the end-user, and it is not unfair to assume
that the average person would not be able to analyze downloaded software for
malware.

% TODO: Add an example of malware distributed via piracy

\subsubsection{Illegal content}
Piracy, as discussed above, provides a way to distribute illegal content. This
content is not always beneficial to society; research articles, art, movies, \&c.
are not the only sorts of content that a pirate may distribute. Piracy opens
the door for questionable and illegal content that may be harmful to society
or national security.

% TODO: Talk about ties to illegal organizations that are involved?

\subsection{Economic}
Piracy, as a theft, also causes real harm to businesses. The act of piracy
does involve stealing and distributing content from creators and organizations
that invested time, money, and effort into the content. In this sense, piracy
is stealing from the individuals' revenue and the government's tax
revenue.~\cite{congress:pirating-the-american-dream}

\textcquote{mrad:effects-of-ip}[.]{%
  According to Gould and Gruben (1996), IPRs protection stimulates economic
  growth if it is accompanied by a policy of trade liberalization. By
  encouraging initiatives to innovate, IPRs protection may influence the
  economic growth of an open country. Park and Ginarte (1997) found that IPRs
  protection affects economic growth indirectly by stimulating the accumulation
  of factors of production such as physical capital and R\&D capital%
}
Piracy, therefore, discourages innovation in the capitalist marketplace.

% TODO
\section{Christian Response}
A Christian response to piracy should honestly take into account both sides of
the debate. As shown above, piracy does have some merit to it. Christians have
to admit that piracy-promoting tools do enable

% 1 Peter 2:13-17
% Romans 13:4-5

\begin{enumerate}
  \item Drawing distinctions: Paywall to information is different from
    withholding information for sake of security

  \item Paywalls provide the means of living for artists, writers, \&c. Providing
    ways to circumvent these when unnecessary may hinder their livelihoods.

  \item Betterment of mankind as a whole, should not be behind a paywall

  \item Piracy provides an avenue for missionary outreach; piracy provides
    means of secret dissemination within an oppressive regime.

  \item Piracy provides the means for access to content/information without
    much accountability
\end{enumerate}

\section{Conclusion}

\subsection{Final Thoughts / Responses}

\begin{enumerate}
  \item In response to economic argument: Would the pirate have bought the
    product/service in the first place?
\end{enumerate}

\begin{displayquote}
  Of additional concern is the fact that by making \textins{Intellectual Property}
  theft a national security threat without being clear about what actually
  constitutes intellectual property, not only does the U.S. government create a
  new reason for a militarized Internet, but it also sets the stage for companies
  to assert that a range of other activities from file sharing to producing
  counterfeit DVDs threaten national security and require further state
  intervention.~\cite[p. 264]{halbert:agendas}
\end{displayquote}

\subsection{Applications}

\subsection{Further Work}

\clearpage
\nocite{*}
\printbibliography
\end{document}

% vim: set ts=2 sw=2 et: