### 2024-04-10 paper.tex
Work on Against and Christian Response sections.
Still need to add more to the Christian response (and include it more
throughout the essay). I also still need to add more to the Against
section, not just the malware section.

\documentclass[onecolumn, 12pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{setspace}

\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{footnote}
\usepackage{endnotes}
\usepackage[bottom, multiple]{footmisc}
\usepackage[level]{fmtcount}

\usepackage{float}
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{ {./graphics} }

\usepackage[style=ieee, backend=biber]{biblatex}
\addbibresource{sources.bib}

\usepackage[pdfusetitle, hidelinks]{hyperref}

\makesavenoteenv{figure}
\SetCiteCommand{\cite}

\title{Sailing the High Seas:\\ A Christian Exploration of Piracy}
\author{Pedro J. Avalos Jim\'enez, Niklas Anderson}
\date{Wheaton College\\\today}

\begin{document}
\maketitle

\section{Introduction}

\subsection{Background Information}

\blockcquote{barlow:wine}{%
  \textins{T}hat ideas should freely spread from one to another over the globe,
  for the moral and mutual instruction of man, and improvement of his
  condition, seems to have been peculiarly and benevolently designed by nature,
  when she made them, like fire, expansible over all space, without lessening
  their density at any point, and like the air in which we breathe, move, and
  have our physical being, incapable of confinement or exclusive appropriation.
  Inventions then cannot, in nature, be a subject of property.
  \emph{Thomas Jefferson}%
}

It is generally accepted that piracy is an illegal activity, but what is often overlooked
is the legitimate role that piracy plays in a variety of contexts. When it comes to
subverting oppressive regimes, for instance, the tools that piracy provides are
indispensable and for protecting the interests of consumers from predatory business
practices, the abilities of piracy are unmatched.

Needless to say, with every unregulated system comes abuse and piracy is no exception.
On top of that, piracy, on an individual level, is extremely difficult to combat because
of the immaterial nature of the items that would need to be protected. Ideas flow freely
between people especially when the Internet is involved, so stopping this flow becomes an
exponentially more difficult problem the more people have the restricted data. The ease
with which ideas propagate lead people like Thomas Jefferson to claim that they are
``incapable of confinement or exclusive appropriation.'' Consequently, two parties form:
those that support piracy for its free and open spread of all information and those that
see piracy as a threat to the further advancement of technology and ideas.

Before proceeding further into the issue of piracy, it is important to have a grasp of
what piracy means insofar as how this paper interprets it. The dictionary defines piracy
as ``the unauthorized use of another's production, invention, or conception especially in
infringement of a copyright.'' While this covers the technical definition of piracy, this
paper elects to utilize a more wholistic definition. Piracy not only includes the
unauthorized use of someone's work, but it includes the tools and techniques that are
commonly used to accomplish this task such as torrent trackers, VPNs, and torrent clients.

As a Christian, how should one approach such a tool and how can it be utilized for the
betterment of God's kingdom? In the paper that follows, piracy will be analyzed through a
number of lenses, culminating in a Christian response to this issue. In doing so, the hope
is that the reader attains a fresh, wholistic understanding of the potential value of
piracy.

\subsection{Thesis/Argument}

Piracy should not be hastily thrown aside.
Piracy is an excellent tool for disseminating information, services, content,
\&c. that may not have been available any other way.

\section{Arguments For Piracy}
As mentioned above, even though piracy is illegal, there is still merit to the tools of
piracy and the values behind them. The merits of piracy can be divided into three realms:
equal access to information, governance, and privacy. 
\subsection{Equal Access}
At its core, piracy is primarily concerned with accessing information of all kinds,
regardless of whether someone is authorized to access that information or not. This is
what piracy is best at and, contrary to popular belief, it is not always a bad thing.

Take the restricted access to academic journal articles, for instance. Many academic
journals require all readers to pay not an insignificant sum to access research articles
that could contain insight that is valuable to the human race as a whole. Because of the
high cost of entry, certain groups are automatically excluded from reading and building
upon the research that has already been done. John Barlow puts the situation eloquently in
\textit{Selling Wine Without Bottles: The Economy of the Mind on the Global Net} when he
says, \textcquote{barlow:wine}{I am not comfortable with a model which will restrict
inquiry to the wealthy.} In the case of academic journal articles, why shouldn't anyone be
able to better themselves by consuming quality research articles? All humans should be
allowed to continue to learn more about God's creation without being bound by their
financial status. Just as how someone can go to the library and checkout any book that
tickles their fancy, they should be freely able to learn from the ongoing research of our
world. Of course one should recognize that not all information is safe or helpful to be
freely available on the internet, but that is a topic for a later section.

A more concrete instance of where paywalls can be directly harmful to the human race is
with regards to medical research. Till et al. stresses how important access to medical
research is in their article \textit{Who is pirating medical literature? A bibliometric
review of 28 million Sci-Hub downloads}. \blockcquote{till:medical-literature} {Access to
the medical literature is essential for both the practice of evidence-based medicine and
meaningful contribution to medical sciences. Nonetheless, only 12\% of newly published
papers are freely accessible online, and, as of 2014, only 3 million of the 26.3 million
articles indexed on PubMed were available on the site's repository of free materials,
PubMed Central. Access to paywall-protected literature remains primarily through
institutional subscriptions. Such subscriptions are costly and many struggle to afford
access. The result is a disparity in access to the medical literature, particularly for
those in low-income and middle-income countries (LMICs).}
According to Till et al. the high subscription cost to access medical journal articles
means that those in less advantaged countries cannot afford access to the material, which
is detrimental to the furthering of medical sciences. In light of this, it should come as
no surprise that \textcquote{till:medical-literature} {Nearly 1 million articles published
by medical journals are downloaded on Sci-Hub each month.} Piracy enables the less
fortunate countries to access valuable medical data that is essential to furthering
medical research for the human race. In such a specific case as this, it should not be
controversial to conclude that piracy is a good thing.

Medical journals are not the only source of information that can sit behind a paywall.
Sometimes, as in the previous case, there is information that one might argue is unjustly
kept away from the common man. What one considers to be unjust is subjective, but the idea
is that piracy is a tool that consumers can employ to fight systems that they deem to be
unjust. Without piracy, one might have to roll over and accept their limited access or
resort to a more extreme and likely illegal approach. To put it another way, piracy acts
as a check on businesses to discourage them from integrating unpopular, anti-consumer
practices into their business model. When a consumer wants to purchase something and they
cannot afford it, they might start to look for alternative way of obtaining it. But, when
they can afford it and they think it is fairly priced, they will purchase it. Thus, piracy
is often where tech-savvy persons turn to when they need something that they feel is
exorbitantly priced, but is not often utilized when they feel the purchase is worthwhile.
This way, piracy acts as a safeguard for consumers when they feel they are being taken
advantage of by predatory business practices.
% XXX the statements in the paragraph above are largely unsubstantiated >:D
% XXX add darnton:pirating-and-publishing and bohannon:everyone sources if needed

\subsection{Governance}
Given that piracy is an illegal usage of another person's conception, it has much to say
in the way of governance and intellectual property laws. The tools of piracy provide the
means of subverting an oppressive regime and call into question the legitimacy of
intellectual property laws.

Having a very firm grip on what citizens can see on the internet is one of the key
characteristics of a despotic government. Such governments know that if they can control
the flow of information on the Internet, it becomes much easier to keep the populace under
control. The government's imposing of restrictions on the Internet is clearly seen when
\textcquote{current:jigsaw} {Tunisia significantly ramped up its already aggressive
blocking of specific websites in response to unrest that would ultimately unseat its
government.} Even more extreme is when governments shutdown the internet entirely to
quickly cease the spread of information. This is exactly what happened on January 28, 2011
in Egypt: \textcquote{current:jigsaw} {Almost simultaneously, about 3,500 individual
Border Gateway Protocol routes \textelp{} were withdrawn on orders from the Egyptian
government, cutting the country off from the rest of the world and bringing internal
communication to a halt.} Fortunately, as already mentioned earlier, piracy excels at
spreading information even when there are those that do not want it spread. BitTorrent is
a peer-to-peer file sharing protocol that is often used when pirating content on the
Internet. A peer-to-peer protocol, by its very nature, does not require a centralized
server to store all the files that are to be shared. Instead, the source of the file is
each person who has the file downloaded. What this means for those under the rule of an
oppressive government is that they can spread information to others in such a way that the
government cannot easily block its source. Pieces of the information are sent to the one
downloading the file from everyone who had the file. There is no single source of the
information that the government can block to stop the spread. By a similar token, there
are peer-to-peer instant messaging applications that do not require an internet connection
to function; you need only be in the vicinity of other people that the message can hop
through to get to the destination. Both of these tools, in the context of piracy, facilitate
the illegal spread of information, but in the context of a despotic government, can be
employed to subvert their attempts at quelling the dispersal of information that could be
harmful to their rule.

Regarding piracy and intellectual property laws, it is clear that the two are at odds with
one another. That being said, it is worth examining whether IP laws accomplish what they
seek to do: encourage innovation. Mark Lemley presents an argument against IP laws when he
writes, \blockcquote[1339]{lemley:faith-based}{[IP] intervenes in the market to interfere
with the freedom of others to do what they want in hopes of achieving the end of
encouraging creativity. If we take that purpose out of the equation, we are left with a
belief system that says the government should restrict your speech and freedom of action
in favor of mine, not because doing so will improve the world, but simply because I spoke
first.} 
To put it differently, intellectual property laws are based on the assumption that in
restricting who is allowed to have certain information, they are encouraging creativity
when, in reality, creativity does not always come first. IP laws are supposed to make
people want to work hard to get a good idea that they can have exclusive ownership of, but
having exclusive rights to an idea stifles further innovation. Under this system, only the
owner has the right to make changes. In fact, it seems that there would be more creativity
without IP laws. If everything was open to scrutiny and replication, there is bound to be
some improvement eventually because of all the people that could have a hand in making it
better. If a big streaming platform were to forfeit one of their IPs to the public, there
is no telling what kind of content could be created. In this way, piracy could allow for
the incremental improvement and scrutiny of conceptions, thus making them better.

\subsection{Privacy}
As a side note, there are instances where pirated software can be more privacy-oriented
than the paid version. Sometimes the pirated version is ``cracked'' to remove certain
digital rights management (DRM) tools that often run background tasks unbeknownst to the
user. These background tasks often leak bits of information about the user in order to
validate that they are legitimate owners of the software. By collecting information about
users, a business invades the privacy of its user base. However, pirated and cracked
software does not have such DRM because it needed to be removed to allow it to be freely
distributed. Thus, when it comes to software that contains aggressive DRM, piracy offers
better security by default. Unfortunately, some pirated software introduces malicious
code, a subject that will be explored in the next section.

\section{Arguments Against Piracy}
While piracy enables free access to information and content, it is crucial to
faithfully and charitably explore the potential dangers and economic
implications associated with this practice. The arguments against piracy are
not easy, nor should be, easy to dismiss. This paper will focus mainly on the
ethical and economical concerns that the use of piracy introduces into society.

\subsection{Access to Dangerous/Illegal Content}
Piracy has been associated with unrestricted access to information and
content. As previously discussed, this aspect of piracy provides the means for
those in need to access information that would otherwise be unavailable to them;
similarly, piracy can thus provide avenues to bypass governmental restrictions
on public information. This same benefit, however, becomes a strong argument
against piracy when what is distributed is dangerous or illegal content.

\subsubsection{Criminality of the Seeder and Leecher}
This paper has focused on a more wholistic definition that encompasses
the tools, techniques, motivations, and content, but the motivations and
content that are involved in piracy should not be ignored. Piracy does involve
a form of theft and dissemination of an individual's or organization's
intellectual property. Piracy, thus can and most often does violate laws in
most jurisdictions. Piracy, when it uses torrenting, involves both the
seeder and the leecher (or the uploader and downloader, if not using torrenting).
Depending on the legislation of the country in question, the leecher may or
may not be found guilty of a crime, but this is complicated by the global
context of digital piracy.

In the United States, the downloader/streamer/leecher are generally considered
offenders that could be subject of fines or even criminal charges. The severity
of the penalties of piracy in these cases pose a major argument against the use
of piracy, at least in the United States or nations where the individual
downloading can be found guilty of a crime. ``Fair use'' and statutes of
limitations may alleviate penalties, but repeat offences may also lose one's
favor in these cases.~\cite{felonies.org}

\subsubsection{Malware}
The intersection of trust and computing is not a new topic. Ken Thompson, the
creator of the Unix operating system, made this the topic of his Turing Award
lecture. In his lecture, he poses the question,
\textcquote[761]{thompson:trust}[.]{%
  to what extent should one trust a statement that
  a program is free of Trojan Horses? Perhaps it is more important to trust the
  people who wrote the software%
} With piracy, however, we add an additional actor in this chain of trust.
Thompson's lecture focuses solely on the developer-to-user relationship of
trust, with the assumption of an open-source project. Thompson, thus, comments
on how \textcquote[763]{thompson:trust}{%
  \textins{s}uch blatant code would not go undetected for long. Even the most
  casual perusal of the source of the C compiler would raise suspicions%
} When piracy is the method of distribution, however, the assumption of
equality between the source code and the delivered software does not hold. The
pirated software may have been tampered with, and the user has no simple way to
detect changes in the file(s) without doing some (often) complicated forensic
work. Although anti-malware tools may mitigate these issues, the problem of
trust still holds, and the anti-malware tools are not perfect. The character of
the seeder must be questioned, and it already stands on shaky ground if the
legality of their actions speaks for their moral compass and trustworthiness.

Piracy may openly distribute dangerous illegal content, but arguably the
larger threat is the dangerous content that remains undisclosed. Piracy
circumvents mainline distribution channels that may incorporate secure policies
of data integrity, authenticity, and privacy. There are piracy tools and sites
that provide some ways to vet the authenticity of a user based on the user's
previous history and relationship with the piracy community. For example,
ThePirateBay, one of the most popular pirating sites, displays the status or
class of their users and holds these users to higher standards in order to
retain trust with the site's end users.~\cite{suprbay:status} Nevertheless, not
all sites provide these metrics or warnings for the average end-user. Instead,
the burden is often on the end-user, and it is not unfair to assume that the
average person would not be able to analyze downloaded software for malware.

The relationship between piracy and malware distribution is not just a
theoretical argument against the use of piracy. Although research has remained
rather theoretical in this area, some researchers have performed empirical
studies to find a stronger link between more prevalent use of piracy and
malware presence. The study by Mezzour et al. reveals that piracy is not equal
around the globe, and it poses serious problems especially for people in
poor countries where piracy is widespread. Although they hypothesized, like
other researchers, that \textcquote{mezzour:empirical-study}[,]{%
  cyber criminals would target rich countries because the monetary benefit of
  compromising rich people's computers is higher
} but they found that, \textcquote{mezzour:empirical-study}[.]{%
  surprisingly, poor countries in Sub-Saharan Africa because of the low cost of
  attacking computers in this region%
} Mezzour et al. conclude that, due to the high correlation between piracy
and malware, one of the policy implications is that Sub-Saharan Africa requires
reducing piracy to combat the spread of malware.

% TODO: Add an example of malware distributed via piracy, see pcmag source

\subsubsection{Illegal content}
Piracy, as discussed above, provides a way to distribute illegal content. This
content is not always beneficial to society; research articles, art, movies, \&c.
are not the only sorts of content that a pirate may distribute. Piracy opens
the door for questionable and illegal content that may be harmful to society
or national security.

% TODO: Talk about ties to illegal organizations that are involved?

\subsection{Economic}
Piracy, as a theft, also causes real harm to businesses. The act of piracy
does involve stealing and distributing content from creators and organizations
that invested time, money, and effort into the content. In this sense, piracy
is stealing from the individuals' revenue and the government's tax
revenue.~\cite{congress:pirating-the-american-dream}

\blockcquote{mrad:effects-of-ip}[.]{%
  According to Gould and Gruben (1996), IPRs protection stimulates economic
  growth if it is accompanied by a policy of trade liberalization. By
  encouraging initiatives to innovate, IPRs protection may influence the
  economic growth of an open country. Park and Ginarte (1997) found that IPRs
  protection affects economic growth indirectly by stimulating the accumulation
  of factors of production such as physical capital and R\&D capital%
}
Piracy, therefore, discourages innovation in the capitalist marketplace.

% TODO
\section{Christian Response}
A Christian response to piracy should honestly take into account both sides of
the debate. As shown above, piracy does have some merit to it. Christians have
to admit that piracy-promoting tools do enable equal access for those in need,
but, at the same time, we must recognize that the human nature is not perfect,
but is full of evil thoughts (Matthew 15:19). Our human corruption manifests
itself in many ways through our lives and actions, but, as explored in this
paper, it also presents itself in the software we share. Our deceitful hearts
(Jeremiah 17:9) make it hard to trust the software we download, but reckless
use of piracy ignores advice and is therefore the way of the fool where people
fall (Proverbs 12:15, 11:14).

On the one hand, a Christian use of piracy should be respectful of authority.
We are called to do this regardless of punishment, but as a matter of
conscience (Romans 13:4--5). We are to exercise our freedom respectfully and
not as a cover-up for evil (1 Peter 2:13--7). In other words, we should consider
both the consequences we may incur and the image we portray carefully before
making a conclusion about the practice of piracy. A proper consideration of
privacy should not ignore the serious ramifications and associations that come
with it.

On the other hand, piracy, when used properly, is able to be used for good
causes that serve God rather than men (Acts 5:29). Academic articles,
especially from medical journals, are crucial for the development of humanity
and the saving of lives. In these cases, when capitalist motivators impede these
greater goods, one should not turn a blind eye, but instead we should remember
the poor and those in need (Proverbs 19:17, 22:9, 15:11; Galatians 2:10). Cases
like these present valid concerns in which piracy appears to provide a viable
solution to the problems brought about by human greed.

\section{Conclusion}

\subsection{Final Thoughts / Responses}

\begin{enumerate}
  \item In response to economic argument: Would the pirate have bought the
    product/service in the first place?
\end{enumerate}

\begin{displayquote}
  Of additional concern is the fact that by making \textins{Intellectual Property}
  theft a national security threat without being clear about what actually
  constitutes intellectual property, not only does the U.S. government create a
  new reason for a militarized Internet, but it also sets the stage for companies
  to assert that a range of other activities from file sharing to producing
  counterfeit DVDs threaten national security and require further state
  intervention.~\cite[p. 264]{halbert:agendas}
\end{displayquote}

\subsection{Applications}

\subsection{Further Work}

\clearpage
\nocite{*}
\printbibliography
\end{document}

% vim: set ts=2 sw=2 et: